Relationships in the extremes and their influence on competition and coexistence

### 2021-09-17 paper/Paper.Rnw
edits to para 4

%***JAS: I am trying to relay the message that climate change and extreme events have consequences for coexistence and competition

@@

Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual increases or decreases over time in an environmental variable. 
Extreme events push environmental variables to their extremes, increasing variance. Effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables. 
For example, decreased growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}. This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity while dominant species become excluded \citep{Mariotte_2013}. Conversely, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}. \cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics. Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.    

>>>>

%***JAS: I am trying to relay the message that climate change and specifically extreme events have consequences for coexistence and competition

@@

Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual shifts over time, whereas increased variance is more likely a product of extreme events as they push environmental variables to their extremes.
Currently, effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables, thus less likely to be a result of extreme events. 
For example, decreased mean growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}. This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity while dominant species become excluded \citep{Mariotte_2013}. Conversely, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}. \cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics. Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as mean rainfall levels are forecasted to decrease.    
